---
title: "Honest model evaluation"
output: rmarkdown::html_vignette
bibliography: references.bib
vignette: >
  %\VignetteIndexEntry{Honest model evaluation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6,
                      fig.height = 4)
```

```{r setup}
library(RDataScienceCompendium)
```

Reported predictive performance is often too optimistic. The usual causes
are information leaking from test data into training, and choosing a model
on the same data used to score it. This vignette shows how to avoid both,
and how to report discrimination *and* calibration.

## Selection bias in cross-validation

We generate an outcome that is pure noise and offer 30 candidate models,
each using a single noise predictor. Every candidate's true out-of-sample
RMSE is about $\mathrm{sd}(y) = 1$ or a little worse, because it is fitting
noise.

```{r noise}
set.seed(1)
n <- 60
p <- 30
noise <- as.data.frame(matrix(rnorm(n * (p + 1)), n))
names(noise) <- c("y", paste0("x", seq_len(p)))

candidates <- lapply(paste0("x", seq_len(p)), function(v) {
  f <- stats::as.formula(paste("y ~", v))
  function(train) lm(f, data = train)
})
names(candidates) <- paste0("x", seq_len(p))
```

The common practice is to cross-validate each candidate, pick the best, and
report its CV score. `nested_cv()` computes that "naive" number alongside
the nested estimate, in which selection is repeated inside every outer
training set [@varma2006; @cawley2010]:

```{r nested}
ncv <- nested_cv(noise, candidates, outcome = "y", outer_k = 5, inner_k = 5,
                 seed = 42)
ncv
```

The naive estimate is lower (better) than the nested one. It is even lower
than the sample standard deviation of $y$, i.e. than the error from simply
predicting the mean. That is the selection bias at work. In the nested
procedure, the outer folds also disagree about which predictor is "best",
a sign that no candidate is really better than the others:

```{r nested-detail}
ncv$outer
```

Averaged over many data sets, the optimism of the naive estimate is
consistently positive; the package tests check this.

## Preventing leakage

`cross_validate()` takes the *whole* modelling procedure as a function. Any
step that learns from data, such as screening predictors, scaling or
imputing, must go inside it so that it is re-learned in each training fold.
Below, the "leaky" version screens predictors on the full data set before
cross-validating. The "honest" version screens inside each fold.

```{r leakage}
screen_fit <- function(train, k = 3) {
  r <- abs(cor(train[-1], train$y))[, 1]
  keep <- names(sort(r, decreasing = TRUE))[seq_len(k)]
  lm(stats::reformulate(keep, "y"), data = train)
}

honest <- cross_validate(noise, screen_fit, outcome = "y", k = 5,
                         repeats = 5, seed = 7)

r_full <- abs(cor(noise[-1], noise$y))[, 1]
keep_full <- names(sort(r_full, decreasing = TRUE))[1:3]
leaky <- cross_validate(
  noise, function(train) lm(stats::reformulate(keep_full, "y"), data = train),
  outcome = "y", k = 5, repeats = 5, seed = 7
)

c(honest = honest$estimate, leaky = leaky$estimate, sd_y = sd(noise$y))
```

Only the honest estimate correctly shows that the procedure does no better
than predicting the mean.

The naive standard error printed by `cross_validate()` treats folds as
independent. It underestimates the real uncertainty and should be read as a
lower bound [@bates2024].

## Discrimination and calibration

For probabilistic classifiers, report a measure of discrimination (such as
the AUC) together with calibration [@vancalster2019] and a proper scoring
rule [@gneiting2007]. We simulate a logistic model with known risks and
compare a correctly specified model with an over-fitted one that has 25
extra noise predictors.

```{r clf}
train <- simulate_logistic(300, beta = c(1, -0.8, 0.5, rep(0, 25)),
                           intercept = -0.5, seed = 11)
test <- simulate_logistic(5000, beta = c(1, -0.8, 0.5, rep(0, 25)),
                          intercept = -0.5, seed = 12)

small <- glm(y ~ x1 + x2 + x3, data = train, family = binomial)
big <- glm(y ~ . - p, data = train, family = binomial)

p_small <- predict(small, test, type = "response")
p_big <- predict(big, test, type = "response")

rbind(
  small = auc_ci(test$y, p_small),
  big = auc_ci(test$y, p_big)
)
```

The over-fitted model loses some discrimination, but its calibration is far
worse:

```{r cal}
rbind(small = calibration(test$y, p_small)[2, ],
      big = calibration(test$y, p_big)[2, ])
c(brier_small = brier_score(test$y, p_small),
  brier_big = brier_score(test$y, p_big),
  logloss_small = log_loss(test$y, p_small),
  logloss_big = log_loss(test$y, p_big))
```

The over-fitted model's calibration slope is well below 1, meaning its
predictions are too extreme. Even the correctly specified model, fitted to
300 observations, has a slope below 1: some over-fitting is unavoidable at
this sample size, which is why shrinkage methods exist. AUC measures only
how well cases are ranked, so it does not reveal how badly the probabilities
are miscalibrated. The Brier score and log loss, which are proper scoring
rules, do.

```{r cal-plot, fig.alt = "Binned calibration plot for the small and big models."}
bin_cal <- function(y, p, bins = 10) {
  g <- cut(p, stats::quantile(p, seq(0, 1, length.out = bins + 1)),
           include.lowest = TRUE)
  data.frame(pred = tapply(p, g, mean), obs = tapply(y, g, mean))
}
cs <- bin_cal(test$y, p_small)
cb <- bin_cal(test$y, p_big)
plot(cs$pred, cs$obs, pch = 19, xlim = c(0, 1), ylim = c(0, 1),
     xlab = "Mean predicted risk", ylab = "Observed event rate")
points(cb$pred, cb$obs, pch = 17, col = 2)
abline(0, 1, lty = 2)
legend("topleft", c("correct model", "over-fitted"), pch = c(19, 17),
       col = 1:2, bty = "n")
```

## References
