---
title: "Resampling inference and effect sizes"
output: rmarkdown::html_vignette
bibliography: references.bib
vignette: >
  %\VignetteIndexEntry{Resampling inference and effect sizes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6,
                      fig.height = 4)
```

```{r setup}
library(RDataScienceCompendium)
```

This vignette shows how to quantify uncertainty without relying on
distributional assumptions (bootstrap intervals and permutation tests), and how
to report the *size* of an effect together with an exact interval.

## Bootstrap confidence intervals

Consider the ratio of mean fuel consumption between automatic and manual cars
in `mtcars`. The ratio has no simple standard error, but the nonparametric
bootstrap gives one directly. For a statistic of a data frame, `boot_ci()`
resamples rows.

```{r boot}
ratio <- function(d) mean(d$mpg[d$am == 0]) / mean(d$mpg[d$am == 1])
ci <- boot_ci(mtcars, ratio, R = 1999, seed = 2024)
ci
```

The four methods rest on different assumptions [@davison1997]:

* **percentile**: quantiles of the bootstrap distribution. It is
  transformation-respecting but only first-order accurate.
* **basic**: reflects the percentile interval around the estimate.
* **normal**: estimate minus bootstrap bias, plus or minus a normal
  quantile times the bootstrap SE.
* **BCa**: corrects the percentile interval for median bias ($\hat z_0$) and
  for skewness via the acceleration $\hat a$ [@efron1987]. It is second-order
  accurate and usually preferred.

```{r boot-diag}
c(bias_correction = ci$bias_correction, acceleration = ci$acceleration)
hist(ci$replicates, breaks = 40, main = "Bootstrap distribution",
     xlab = "ratio of means", col = "grey85", border = "white")
abline(v = ci$estimate, lwd = 2)
```

`boot_ci()` reproduces `boot::boot.ci()` to machine precision given the same
seed (see the package tests). What it adds is a small, validated interface and
a guarantee that the caller's random number stream is not changed.

## Permutation tests

A permutation test gives an exact test of the null hypothesis that the group
labels are exchangeable. With 13 manual and 19 automatic cars there are
$\binom{32}{13} \approx 3.5 \times 10^{8}$ rearrangements, far too many to
enumerate, so `perm_test()` draws a random sample of them:

```{r perm}
manual <- mtcars$mpg[mtcars$am == 1]
auto <- mtcars$mpg[mtcars$am == 0]
pt <- perm_test(manual, auto, R = 9999, seed = 1)
pt
pt$p.value.mcse
```

The p-value is computed as $(b + 1)/(R + 1)$, never as $b/R$. The
$b/R$ estimator can return zero, and the test built on it is anti-conservative
[@phipson2010]. Any statistic can be used, such as a difference in medians:

```{r perm-median}
perm_test(manual, auto, statistic = function(x, y) median(x) - median(y),
          R = 9999, seed = 1)$p.value
```

When the reference set is small, the test is exact. In the paired `sleep`
data there are $2^{10} = 1024$ sign flips:

```{r perm-exact}
perm_test(sleep$extra[1:10], sleep$extra[11:20], paired = TRUE)
```

## Effect sizes with exact intervals

A p-value says nothing about the magnitude of an effect. Standardised mean
differences do, but their commonly reported large-sample intervals can be
badly wrong in small samples. `cohens_d()` and `hedges_g()` invert the
noncentral $t$ distribution instead [@steiger1997; @cumming2001]:

```{r es}
hedges_g(manual, auto)
cohens_d(sleep$extra[1:10], sleep$extra[11:20], paired = TRUE)
```

Hedges' $g$ applies the exact correction $J(\nu)$, which removes the upward
bias of $d$ in small samples [@hedges1981]:

```{r hedges-j}
round(hedges_correction(c(5, 10, 20, 50, 100)), 4)
```

## References
