---
title: "Designing and reporting simulation studies"
output: rmarkdown::html_vignette
bibliography: references.bib
vignette: >
  %\VignetteIndexEntry{Designing and reporting simulation studies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6.5,
                      fig.height = 4)
```

```{r setup}
library(RDataScienceCompendium)
```

Simulation studies are how statisticians find out whether a method works. Like
any experiment, they need a pre-specified design, an adequate number of
replicates, and uncertainty statements about their results. This vignette
follows the ADEMP structure of @morris2019 and runs a small study. The
full-size version, with more replicates and scenarios, is in the `analysis/`
directory of the repository.

## A: Aims

To compare the coverage of four 95% confidence intervals for a population mean
when the data are right-skewed and the sample is small.

## D: Data-generating mechanisms

Samples of size $n \in \{10, 40\}$ from a log-normal distribution with
$\log X \sim N(0, \sigma^2)$, with $\sigma \in \{0.5, 1\}$ (moderate and
strong skew). That is a $2 \times 2$ factorial design.

```{r dgm}
scenarios <- expand.grid(n = c(10, 40), sigma = c(0.5, 1))
generate <- function(n, sigma) rlnorm(n, meanlog = 0, sdlog = sigma)
scenarios
```

## E: Estimand

The population mean $\theta = \exp(\sigma^2/2)$. It varies by scenario, so we
store it as a column.

## M: Methods

1. The Student $t$ interval.
2. The bootstrap percentile interval.
3. The bootstrap BCa interval.
4. The bootstrap basic interval.

The analysis function returns one row per method:

```{r methods}
analyse <- function(x) {
  tt <- stats::t.test(x)
  bs <- boot_ci(x, mean, R = 499, type = c("percentile", "bca", "basic"))
  data.frame(
    method = c("t", bs$intervals$type),
    estimate = mean(x),
    se = c(sd(x) / sqrt(length(x)), rep(bs$se, 3)),
    lower = c(tt$conf.int[1], bs$intervals$lower),
    upper = c(tt$conf.int[2], bs$intervals$upper)
  )
}
```

## P: Performance measures and number of replicates

The key measure is **coverage**. To estimate a coverage near 95% with a Monte
Carlo SE of 1 percentage point, we need

```{r nsim}
sim_n_required(0.01, p = 0.95)
```

replicates per scenario. To keep this vignette fast we use 500, which gives
MCSE $\approx 1$ point. The study in `analysis/` uses 5,000 (MCSE
$\approx 0.3$ points).

## Running the study

```{r run}
res <- run_simulation(generate, analyse, n_sim = 500, scenarios = scenarios,
                      seed = 20240901)
res$theta <- exp(res$sigma^2 / 2)
res
```

Every replicate gets its own L'Ecuyer-CMRG stream [@lecuyer2002]. The results
are therefore identical if the study is re-run in parallel, e.g. with
`future::plan("multisession")` and `backend = "future"`.

## Results

```{r perf}
perf <- sim_performance(res, true = "theta", by = c("n", "sigma", "method"))
cov <- subset(perf, measure == "coverage")
cov[, c("n", "sigma", "method", "estimate", "mcse")]
```

Coverage estimates should always be plotted with their Monte Carlo
uncertainty:

```{r plot, fig.alt = "Coverage of four interval methods by scenario, with Monte Carlo error bars."}
cov$label <- sprintf("n=%d, sigma=%.1f", cov$n, cov$sigma)
methods <- unique(cov$method)
labels <- unique(cov$label)
pos <- outer(seq_along(labels), (seq_along(methods) - 2.5) * 0.18, "+")
plot(NA, xlim = c(0.5, length(labels) + 0.5), ylim = c(0.75, 1),
     xaxt = "n", xlab = "", ylab = "Coverage")
axis(1, at = seq_along(labels), labels = labels, cex.axis = 0.8)
abline(h = 0.95, lty = 2)
for (m in seq_along(methods)) {
  d <- cov[cov$method == methods[m], ]
  x <- pos[match(d$label, labels), m]
  arrows(x, d$estimate - 1.96 * d$mcse, x, d$estimate + 1.96 * d$mcse,
         angle = 90, code = 3, length = 0.02, col = m)
  points(x, d$estimate, pch = 19, col = m)
}
legend("bottomright", legend = methods, col = seq_along(methods), pch = 19,
       bty = "n", cex = 0.8)
```

Three findings stand out. (i) All methods under-cover when the skew is
strong, and badly so when $n = 10$: bootstrap intervals for a mean are too
narrow in small samples, partly because the plug-in variance divides by $n$
rather than $n - 1$ [@davison1997]. (ii) The basic interval is clearly the
worst, as theory predicts for skewed statistics, because it reflects the
bootstrap distribution the wrong way. (iii) With 500 replicates, the
differences between the $t$, percentile and BCa intervals are mostly
comparable to their Monte Carlo SEs, so this small study **cannot** rank
them. Without MCSEs it would be tempting to over-interpret those
differences. The larger study in `analysis/` is designed to resolve them.

### Other performance measures

`sim_performance()` also reports bias, empirical and model-based SEs, MSE and
interval width:

```{r other}
subset(perf, method == "t" & measure %in% c("bias", "empirical_se",
                                             "rel_error_model_se"))
```

## Reporting checklist

Following @morris2019, a report should state:

* the aims, data-generating mechanisms, estimands, methods and performance
  measures (ADEMP);
* the number of replicates and how it was chosen;
* the random number generator and seed (`attr(res, "seed")`);
* how many replicates failed (`sum(!is.na(res$.error))`);
* Monte Carlo SEs for every performance estimate.

## References
