---
title: "Business Intelligence Case Study: E-commerce Analytics"
subtitle: "Advanced R Analytics for Customer Segmentation and Revenue Optimization"
author: "R Data Science Portfolio"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: flatly
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    number_sections: true
    code_folding: show
    df_print: paged
    fig_width: 12
    fig_height: 8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  warning = FALSE, 
  message = FALSE,
  fig.align = "center",
  cache = FALSE,
  dpi = 300
)

# Fixed analysis date so that results do not depend on the render date
analysis_date <- as.Date("2024-12-31")

# Load required libraries
library(dplyr)
library(ggplot2)
library(plotly)
library(DT)
library(kableExtra)
library(corrplot)
library(cluster)
library(factoextra)
library(lubridate)
library(scales)
library(viridis)
library(patchwork)
library(tidyr)
library(stringr)
library(forcats)
library(readr)

# Source custom functions (adjust path as needed)
if (file.exists("../utils/data-generators.R")) {
  source("../utils/data-generators.R")
}
if (file.exists("../03-visualization/ggplot2-mastery.R")) {
  source("../03-visualization/ggplot2-mastery.R")
}

# Set theme for all plots
theme_set(theme_minimal() + 
          theme(plot.title = element_text(size = 14, face = "bold"),
                plot.subtitle = element_text(size = 12, color = "gray40")))
```

# Executive Summary {.tabset}

## Key Findings

This case study walks through a customer-analytics workflow for an e-commerce
business: data quality checks, exploratory analysis, RFM and k-means
segmentation, CLV and churn-risk scoring, and scenario analysis.

**All data are simulated** by `generate_comprehensive_business_data()` in
`modules/utils/data-generators.R`, with a fixed seed and a fixed analysis date
(`r analysis_date`). The findings therefore illustrate the methods, not a real
company. Every number quoted in the [conclusion](#conclusion) is computed from
the analysis below.

## Methodology

Our analysis employs a comprehensive data science approach:

1. **Data Generation & Validation**: Simulated customers, products and transactions (sizes in the *Dataset Generation* chunk)
2. **Exploratory Data Analysis**: Statistical profiling and pattern identification
3. **Customer Segmentation**: RFM scoring and k-means clustering, with the number of clusters chosen by average silhouette width
4. **Customer Scoring**: Rule-based CLV and churn-risk scores
5. **Scenario Analysis**: Revenue sensitivity to explicitly stated uplift assumptions

---

# Data Overview & Quality Assessment

## Dataset Generation

```{r data-generation}
# Generate comprehensive business dataset
set.seed(42)
business_data <- generate_comprehensive_business_data(
  n_customers = 1000,
  n_products = 150,
  n_transactions = 2500,
  date_range = c(as.Date("2023-01-01"), analysis_date)
)

# Extract main datasets
customers <- business_data$customers
products <- business_data$products
transactions <- business_data$transactions
customer_metrics <- business_data$customer_metrics
```

## Data Quality Assessment

```{r data-quality}
# Comprehensive data quality report
quality_report <- list()

# Customer data quality
quality_report$customers <- customers %>%
  summarise(
    total_records = n(),
    missing_data = sum(is.na(.)),
    duplicate_customers = sum(duplicated(customer_id)),
    email_validity = sum(str_detect(email, "^[\\w\\.-]+@[\\w\\.-]+\\.[a-zA-Z]{2,}$")) / n() * 100,
    age_range = paste(min(age, na.rm = TRUE), "-", max(age, na.rm = TRUE)),
    income_range = paste("$", scales::comma(min(annual_income, na.rm = TRUE)), 
                        " - $", scales::comma(max(annual_income, na.rm = TRUE)))
  )

# Transaction data quality
quality_report$transactions <- transactions %>%
  summarise(
    total_transactions = n(),
    completed_transactions = sum(status == "Completed"),
    completion_rate = round(sum(status == "Completed") / n() * 100, 1),
    avg_order_value = round(mean(total_amount[status == "Completed"], na.rm = TRUE), 2),
    total_revenue = scales::dollar(sum(total_amount[status == "Completed"], na.rm = TRUE)),
    date_range = paste(min(transaction_date), "to", max(transaction_date))
  )

# Display quality metrics
quality_report %>%
  map_dfr(~as.data.frame(t(.)), .id = "Dataset") %>%
  kbl(caption = "Data Quality Assessment") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

## Dataset Overview

```{r dataset-overview}
# Create comprehensive dataset summary
dataset_summary <- tribble(
  ~Dataset, ~Records, ~Variables, ~Description,
  "Customers", nrow(customers), ncol(customers), "Customer demographics, segments, and behaviors",
  "Products", nrow(products), ncol(products), "Product catalog with pricing and categories", 
  "Transactions", nrow(transactions), ncol(transactions), "Purchase history and transaction details",
  "Customer Metrics", nrow(customer_metrics), ncol(customer_metrics), "Derived customer performance indicators"
)

dataset_summary %>%
  kbl(caption = "Dataset Structure Overview") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#3498DB")
```

---

# Exploratory Data Analysis {.tabset}

## Customer Demographics

```{r customer-demographics, fig.height=10}
# Customer demographic analysis
demo_plots <- list()

# Age distribution by segment
demo_plots$age <- ggplot(customers, aes(x = age, fill = customer_segment)) +
  geom_histogram(bins = 25, alpha = 0.8, position = "stack") +
  scale_fill_manual(values = get_custom_palette("professional", 4)) +
  labs(title = "Customer Age Distribution by Segment",
       x = "Age", y = "Count", fill = "Segment") +
  theme_professional()

# Income distribution
demo_plots$income <- ggplot(customers, aes(x = customer_segment, y = annual_income, fill = customer_segment)) +
  geom_boxplot(alpha = 0.8, outlier.alpha = 0.6) +
  geom_jitter(width = 0.2, alpha = 0.3, size = 0.8) +
  scale_fill_manual(values = get_custom_palette("professional", 4)) +
  scale_y_continuous(labels = dollar_format()) +
  labs(title = "Income Distribution by Customer Segment",
       x = "Customer Segment", y = "Annual Income") +
  theme_professional() +
  theme(legend.position = "none")

# Geographic distribution
geo_summary <- customers %>%
  count(country, customer_segment) %>%
  group_by(country) %>%
  mutate(total = sum(n), pct = n/total * 100)

demo_plots$geography <- ggplot(geo_summary, aes(x = reorder(country, total), y = n, fill = customer_segment)) +
  geom_col(position = "stack", alpha = 0.8) +
  scale_fill_manual(values = get_custom_palette("professional", 4)) +
  labs(title = "Customer Distribution by Country and Segment",
       x = "Country", y = "Number of Customers", fill = "Segment") +
  theme_professional() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Combine plots
combined_demo <- (demo_plots$age / demo_plots$income) | demo_plots$geography
combined_demo + plot_annotation(
  title = "Customer Demographics Analysis",
  subtitle = "Understanding our customer base composition and characteristics"
)
```

### Key Demographic Insights

```{r demographic-insights}
# Calculate key demographic statistics
demo_insights <- customers %>%
  group_by(customer_segment) %>%
  summarise(
    avg_age = round(mean(age, na.rm = TRUE), 1),
    avg_income = scales::dollar(round(mean(annual_income, na.rm = TRUE), 0)),
    avg_satisfaction = round(mean(satisfaction_score, na.rm = TRUE), 2),
    pct_active = round(sum(is_active) / n() * 100, 1),
    .groups = "drop"
  ) %>%
  arrange(desc(avg_age))

demo_insights %>%
  kbl(col.names = c("Segment", "Avg Age", "Avg Income", "Satisfaction", "% Active"),
      caption = "Customer Segment Characteristics") %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  row_spec(which(demo_insights$customer_segment == "Premium"), 
           bold = TRUE, background = "#e8f5e8")
```

## Purchase Behavior

```{r purchase-behavior, fig.height=10}
# Analyze purchase patterns
purchase_data <- transactions %>%
  filter(status == "Completed") %>%
  left_join(customers %>% select(customer_id, customer_segment), by = "customer_id")

behavior_plots <- list()

# Order value distribution
behavior_plots$order_value <- ggplot(purchase_data, aes(x = total_amount, fill = customer_segment)) +
  geom_histogram(bins = 50, alpha = 0.7, position = "identity") +
  scale_fill_manual(values = get_custom_palette("professional", 4)) +
  scale_x_continuous(labels = dollar_format(), limits = c(0, 1000)) +
  facet_wrap(~customer_segment, scales = "free_y") +
  labs(title = "Order Value Distribution by Customer Segment",
       x = "Order Value ($)", y = "Frequency") +
  theme_professional() +
  theme(legend.position = "none")

# Channel preferences
channel_data <- purchase_data %>%
  count(customer_segment, channel) %>%
  group_by(customer_segment) %>%
  mutate(pct = n / sum(n) * 100)

behavior_plots$channel <- ggplot(channel_data, aes(x = customer_segment, y = pct, fill = channel)) +
  geom_col(position = "dodge", alpha = 0.8) +
  scale_fill_manual(values = get_custom_palette("business", 4)) +
  labs(title = "Channel Preferences by Customer Segment",
       x = "Customer Segment", y = "Percentage of Purchases", fill = "Channel") +
  theme_professional()

# Monthly purchase patterns
monthly_patterns <- purchase_data %>%
  mutate(month = month(transaction_date, label = TRUE)) %>%
  group_by(month, customer_segment) %>%
  summarise(
    total_revenue = sum(total_amount),
    avg_order_value = mean(total_amount),
    .groups = "drop"
  )

behavior_plots$monthly <- ggplot(monthly_patterns, aes(x = month, y = total_revenue, 
                                                      color = customer_segment, group = customer_segment)) +
  geom_line(size = 1.2, alpha = 0.8) +
  geom_point(size = 3, alpha = 0.8) +
  scale_color_manual(values = get_custom_palette("professional", 4)) +
  scale_y_continuous(labels = dollar_format()) +
  labs(title = "Monthly Revenue Patterns by Customer Segment",
       x = "Month", y = "Total Revenue", color = "Segment") +
  theme_professional()

# Combine plots
(behavior_plots$order_value / behavior_plots$channel) | behavior_plots$monthly
```

### Purchase Behavior Summary

```{r purchase-summary}
# Calculate purchase behavior metrics
purchase_summary <- purchase_data %>%
  group_by(customer_segment) %>%
  summarise(
    avg_order_value = dollar(mean(total_amount)),
    median_order_value = dollar(median(total_amount)),
    total_orders = scales::comma(n()),
    total_revenue = dollar(sum(total_amount)),
    avg_discount = paste0(round(mean(discount_percent), 1), "%"),
    preferred_channel = names(sort(table(channel), decreasing = TRUE))[1],
    .groups = "drop"
  )

purchase_summary %>%
  kbl(caption = "Purchase Behavior by Customer Segment") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  row_spec(which(purchase_summary$customer_segment == "Premium"), 
           bold = TRUE, background = "#fff3cd")
```

## Product Performance

```{r product-performance, fig.height=8}
# Product performance analysis
product_performance <- transactions %>%
  filter(status == "Completed") %>%
  left_join(products %>% select(product_id, category, price, avg_rating), by = "product_id") %>%
  group_by(category) %>%
  summarise(
    total_revenue = sum(total_amount),
    units_sold = sum(quantity),
    avg_price = mean(price),
    avg_rating = mean(avg_rating),
    unique_customers = n_distinct(customer_id),
    .groups = "drop"
  ) %>%
  arrange(desc(total_revenue))

# Revenue by category
p1 <- ggplot(product_performance, aes(x = reorder(category, total_revenue), y = total_revenue)) +
  geom_col(fill = "#3498DB", alpha = 0.8) +
  coord_flip() +
  scale_y_continuous(labels = dollar_format()) +
  labs(title = "Revenue by Product Category",
       x = "Category", y = "Total Revenue") +
  theme_professional()

# Price vs Rating scatter
p2 <- ggplot(product_performance, aes(x = avg_price, y = avg_rating, size = total_revenue)) +
  geom_point(alpha = 0.7, color = "#E74C3C") +
  scale_size_continuous(labels = dollar_format(), name = "Revenue") +
  scale_x_continuous(labels = dollar_format()) +
  labs(title = "Price vs Rating Analysis",
       x = "Average Price", y = "Average Rating") +
  theme_professional()

p1 + p2
```

---

# Advanced Customer Segmentation {.tabset}

## RFM Analysis

```{r rfm-analysis}
# Calculate RFM metrics (Recency, Frequency, Monetary)
reference_date <- max(transactions$transaction_date) + 1

rfm_data <- transactions %>%
  filter(status == "Completed") %>%
  group_by(customer_id) %>%
  summarise(
    last_purchase_date = max(transaction_date),
    frequency = n(),
    monetary = sum(total_amount),
    .groups = "drop"
  ) %>%
  mutate(
    recency = as.numeric(reference_date - last_purchase_date),
    # Create quintiles for each metric
    R_score = ntile(desc(recency), 5),  # Lower recency = higher score
    F_score = ntile(frequency, 5),
    M_score = ntile(monetary, 5),
    # Combined RFM score
    RFM_score = paste0(R_score, F_score, M_score),
    # Customer value segments
    customer_value_segment = case_when(
      R_score >= 4 & F_score >= 4 & M_score >= 4 ~ "Champions",
      R_score >= 3 & F_score >= 3 & M_score >= 3 ~ "Loyal Customers", 
      R_score >= 3 & F_score <= 2 & M_score >= 4 ~ "Big Spenders",
      R_score <= 2 & F_score >= 4 ~ "At Risk",
      R_score <= 2 & F_score <= 2 & M_score <= 2 ~ "Lost Customers",
      TRUE ~ "Others"
    )
  )

# RFM segment summary
rfm_summary <- rfm_data %>%
  group_by(customer_value_segment) %>%
  summarise(
    customer_count = n(),
    avg_recency = round(mean(recency), 1),
    avg_frequency = round(mean(frequency), 1),
    avg_monetary = dollar(round(mean(monetary), 0)),
    total_value = dollar(sum(monetary)),
    .groups = "drop"
  ) %>%
  arrange(desc(parse_number(total_value)))

rfm_summary %>%
  kbl(caption = "RFM Customer Segmentation Analysis") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  row_spec(1, bold = TRUE, background = "#d4edda") %>%
  row_spec(nrow(rfm_summary), color = "red", background = "#f8d7da")
```

## K-Means Clustering

```{r kmeans-clustering, fig.height=10}
# Prepare data for clustering. Use one set of complete-case rows for both the
# clustering matrix and the profile table so that labels stay aligned.
cluster_vars <- c("total_spent", "avg_order_value", "purchase_frequency",
                  "age", "annual_income", "satisfaction_score")
cluster_input <- customer_metrics %>%
  left_join(customers %>% select(customer_id, age, annual_income, satisfaction_score,
                                 customer_segment),
            by = "customer_id") %>%
  filter(if_all(all_of(cluster_vars), ~ !is.na(.x)))
clustering_data <- scale(as.matrix(cluster_input[cluster_vars]))

# Elbow plot for reference
set.seed(42)
fviz_nbclust(clustering_data, kmeans, method = "wss", k.max = 10) +
  labs(title = "Elbow Method for Optimal Clusters") +
  theme_professional()

# Choose k by the average silhouette width (Rousseeuw, 1987)
dist_matrix <- dist(clustering_data)
k_candidates <- 2:8
avg_silhouette <- sapply(k_candidates, function(kk) {
  set.seed(42)
  cl <- kmeans(clustering_data, centers = kk, nstart = 25)$cluster
  mean(cluster::silhouette(cl, dist_matrix)[, "sil_width"])
})
k <- k_candidates[which.max(avg_silhouette)]
cat("Average silhouette width by k:",
    paste(sprintf("k=%d: %.3f", k_candidates, avg_silhouette), collapse = ", "),
    "\nSelected k =", k, "\n")

# Perform k-means clustering
set.seed(42)
kmeans_result <- kmeans(clustering_data, centers = k, nstart = 25)

# Add cluster assignments
cluster_analysis <- cluster_input %>%
  mutate(cluster = factor(kmeans_result$cluster))

# Visualize clusters
cluster_plots <- list()

# Cluster visualization (first two principal components)
pca_result <- prcomp(clustering_data, scale = FALSE)
pca_data <- data.frame(pca_result$x[,1:2]) %>%
  mutate(cluster = factor(kmeans_result$cluster))

cluster_plots$pca <- ggplot(pca_data, aes(x = PC1, y = PC2, color = cluster)) +
  geom_point(alpha = 0.7, size = 2) +
  scale_color_manual(values = get_custom_palette("professional", k)) +
  labs(title = "Customer Clusters (PCA Visualization)",
       x = "First Principal Component", 
       y = "Second Principal Component") +
  theme_professional()

# Cluster characteristics
cluster_plots$characteristics <- cluster_analysis %>%
  group_by(cluster) %>%
  summarise(
    count = n(),
    avg_spent = mean(total_spent, na.rm = TRUE),
    avg_frequency = mean(purchase_frequency, na.rm = TRUE),
    avg_income = mean(annual_income, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  gather(metric, value, -cluster, -count) %>%
  ggplot(aes(x = cluster, y = value, fill = cluster)) +
  geom_col(alpha = 0.8) +
  facet_wrap(~metric, scales = "free_y") +
  scale_fill_manual(values = get_custom_palette("professional", k)) +
  labs(title = "Cluster Characteristics Comparison",
       x = "Cluster", y = "Average Value") +
  theme_professional() +
  theme(legend.position = "none")

cluster_plots$pca / cluster_plots$characteristics
```

## Segment Profiling

```{r segment-profiling}
# Detailed cluster profiling
cluster_profiles <- cluster_analysis %>%
  group_by(cluster) %>%
  summarise(
    customers = n(),
    avg_total_spent = dollar(round(mean(total_spent, na.rm = TRUE), 0)),
    avg_order_value = dollar(round(mean(avg_order_value, na.rm = TRUE), 0)),
    avg_frequency = round(mean(purchase_frequency, na.rm = TRUE), 2),
    avg_age = round(mean(age, na.rm = TRUE), 1),
    avg_income = dollar(round(mean(annual_income, na.rm = TRUE), 0)),
    avg_satisfaction = round(mean(satisfaction_score, na.rm = TRUE), 2),
    dominant_segment = names(sort(table(customer_segment), decreasing = TRUE))[1],
    .groups = "drop"
  ) %>%
  arrange(desc(parse_number(avg_total_spent)))

# Add strategic recommendations
cluster_profiles <- cluster_profiles %>%
  mutate(
    strategy = case_when(
      cluster == cluster_profiles$cluster[1] ~ "VIP Treatment - Maximize retention",
      cluster == cluster_profiles$cluster[2] ~ "Growth Focus - Upsell opportunities", 
      cluster == cluster_profiles$cluster[3] ~ "Engagement - Increase purchase frequency",
      TRUE ~ "Reactivation - Win-back campaigns"
    )
  )

cluster_profiles %>%
  kbl(caption = "Detailed Customer Cluster Profiles with Strategic Recommendations") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  column_spec(ncol(cluster_profiles), background = "#e8f4fd", bold = TRUE) %>%
  row_spec(1, background = "#d4edda")
```

---

# Predictive Analytics {.tabset}

## Customer Lifetime Value Prediction

```{r clv-prediction}
# Calculate and predict Customer Lifetime Value
clv_data <- customer_metrics %>%
  left_join(customers %>% select(customer_id, age, customer_segment, annual_income, registration_date),
            by = "customer_id") %>%
  mutate(
    customer_age_days = as.numeric(analysis_date - registration_date),
    # Historical CLV calculation
    historical_clv = total_spent,
    # Predicted CLV using multiple factors
    predicted_clv = case_when(
      customer_segment == "Premium" ~ total_spent * 2.5 + (purchase_frequency * 200),
      customer_segment == "Gold" ~ total_spent * 2.0 + (purchase_frequency * 150),
      customer_segment == "Standard" ~ total_spent * 1.5 + (purchase_frequency * 100),
      TRUE ~ total_spent * 1.2 + (purchase_frequency * 75)
    ),
    # CLV category
    clv_category = case_when(
      predicted_clv >= 2000 ~ "High Value",
      predicted_clv >= 1000 ~ "Medium Value", 
      predicted_clv >= 500 ~ "Low Value",
      TRUE ~ "Very Low Value"
    )
  )

# CLV distribution
clv_plot <- ggplot(clv_data, aes(x = predicted_clv, fill = customer_segment)) +
  geom_histogram(bins = 30, alpha = 0.7, position = "identity") +
  scale_fill_manual(values = get_custom_palette("professional", 4)) +
  scale_x_continuous(labels = dollar_format()) +
  labs(title = "Predicted Customer Lifetime Value Distribution",
       x = "Predicted CLV ($)", y = "Number of Customers", fill = "Segment") +
  theme_professional()

# CLV vs actual spending
clv_correlation <- ggplot(clv_data, aes(x = total_spent, y = predicted_clv, color = customer_segment)) +
  geom_point(alpha = 0.7, size = 2) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_manual(values = get_custom_palette("professional", 4)) +
  scale_x_continuous(labels = dollar_format()) +
  scale_y_continuous(labels = dollar_format()) +
  labs(title = "CLV Prediction vs Historical Spending",
       x = "Historical Spending ($)", y = "Predicted CLV ($)", color = "Segment") +
  theme_professional()

clv_plot / clv_correlation
```

### CLV Analysis Summary

```{r clv-summary}
# CLV summary statistics
clv_summary <- clv_data %>%
  group_by(customer_segment) %>%
  summarise(
    customers = n(),
    avg_historical_clv = dollar(round(mean(historical_clv, na.rm = TRUE), 0)),
    avg_predicted_clv = dollar(round(mean(predicted_clv, na.rm = TRUE), 0)),
    clv_uplift = round((mean(predicted_clv, na.rm = TRUE) / mean(historical_clv, na.rm = TRUE) - 1) * 100, 1),
    high_value_customers = sum(clv_category == "High Value"),
    .groups = "drop"
  ) %>%
  arrange(desc(parse_number(avg_predicted_clv)))

clv_summary %>%
  mutate(clv_uplift = paste0(clv_uplift, "%")) %>%
  kbl(caption = "Customer Lifetime Value Analysis by Segment") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  column_spec(4, background = "#e8f5e8", bold = TRUE)
```

## Churn Risk Assessment

```{r churn-analysis}
# Calculate churn risk indicators
churn_data <- customer_metrics %>%
  left_join(customers %>% select(customer_id, customer_segment, registration_date),
            by = "customer_id") %>%
  mutate(
    days_since_last_purchase = as.numeric(analysis_date - last_purchase),
    customer_lifetime_days = as.numeric(analysis_date - registration_date),
    # Churn risk factors
    recency_risk = case_when(
      days_since_last_purchase > 90 ~ 3,
      days_since_last_purchase > 60 ~ 2,
      days_since_last_purchase > 30 ~ 1,
      TRUE ~ 0
    ),
    frequency_risk = case_when(
      purchase_frequency < 0.5 ~ 3,  # Less than 0.5 purchases per month
      purchase_frequency < 1.0 ~ 2,
      purchase_frequency < 2.0 ~ 1,
      TRUE ~ 0
    ),
    monetary_risk = case_when(
      avg_order_value < 50 ~ 3,
      avg_order_value < 100 ~ 2,
      avg_order_value < 200 ~ 1,
      TRUE ~ 0
    ),
    # Combined churn risk score
    churn_risk_score = recency_risk + frequency_risk + monetary_risk,
    churn_risk_level = case_when(
      churn_risk_score >= 7 ~ "High Risk",
      churn_risk_score >= 4 ~ "Medium Risk",
      churn_risk_score >= 2 ~ "Low Risk",
      TRUE ~ "Healthy"
    )
  )

# Churn risk distribution
churn_distribution <- churn_data %>%
  count(churn_risk_level, customer_segment) %>%
  group_by(customer_segment) %>%
  mutate(pct = n / sum(n) * 100)

churn_plot <- ggplot(churn_distribution, aes(x = customer_segment, y = pct, fill = churn_risk_level)) +
  geom_col(position = "stack", alpha = 0.8) +
  scale_fill_manual(values = c("Healthy" = "#27AE60", "Low Risk" = "#F39C12", 
                              "Medium Risk" = "#E67E22", "High Risk" = "#E74C3C")) +
  labs(title = "Churn Risk Distribution by Customer Segment",
       x = "Customer Segment", y = "Percentage", fill = "Churn Risk") +
  theme_professional()

print(churn_plot)

# High-risk customer analysis
high_risk_summary <- churn_data %>%
  filter(churn_risk_level == "High Risk") %>%
  group_by(customer_segment) %>%
  summarise(
    high_risk_customers = n(),
    avg_days_since_purchase = round(mean(days_since_last_purchase), 0),
    avg_total_spent = dollar(round(mean(total_spent), 0)),
    potential_lost_revenue = dollar(round(sum(total_spent), 0)),
    .groups = "drop"
  )

high_risk_summary %>%
  kbl(caption = "High-Risk Customer Analysis - Immediate Action Required") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#E74C3C")
```

---

# Business Recommendations {.tabset}

## Strategic Initiatives

### 🎯 Primary Recommendations

The analysis suggests three initiatives. Budgets and expected returns are not given: they would depend on real costs and on intervention effects, which have to be measured (for example in A/B tests), not assumed.

#### 1. Premium Customer Retention Program
- **Target**: Premium segment customers (highest CLV)
- **Actions**: 
  - Dedicated account management
  - Exclusive product previews
  - Personalized offers based on purchase history

#### 2. Mid-Tier Customer Activation
- **Target**: Standard segment customers with growth potential
- **Actions**:
  - Targeted upselling campaigns
  - Category-specific recommendations
  - Loyalty point acceleration programs

#### 3. Churn Prevention Campaign
- **Target**: High-risk customers (`r scales::percent(mean(churn_data$churn_risk_level == "High Risk"), accuracy = 0.1)` of customers with purchases)
- **Actions**:
  - Win-back email sequences
  - Special discount offers
  - Personal outreach for high-value at-risk customers

### 📊 Implementation Roadmap

```{r implementation-roadmap}
# Create implementation timeline
roadmap <- tribble(
  ~Initiative, ~Phase, ~Timeline, ~Success_Measure,
  "Premium Retention", "Phase 1", "Quarter 1", "Retention of premium customers vs. hold-out group",
  "Premium Retention", "Phase 2", "Quarter 2", "Repeat-purchase rate vs. hold-out group",
  "Mid-Tier Activation", "Phase 1", "Quarter 2", "Order frequency vs. hold-out group",
  "Mid-Tier Activation", "Phase 2", "Quarter 3", "Average order value vs. hold-out group",
  "Churn Prevention", "Phase 1", "Quarter 1", "Reactivation rate of high-risk customers",
  "Churn Prevention", "Phase 2", "Quarter 2", "Share of customers moving out of high risk"
)

roadmap %>%
  kbl(caption = "Strategic Implementation Roadmap") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  pack_rows("Premium Customer Focus", 1, 2, background = "#fff3cd") %>%
  pack_rows("Growth Initiatives", 3, 4, background = "#e8f4fd") %>%
  pack_rows("Retention Programs", 5, 6, background = "#f8d7da")
```

## Scenario Analysis

The simulated data cannot tell us how much an intervention would change
customer behaviour. So instead of predicting financial impact, this section
shows how annual revenue would respond to a **range of assumed uplifts**,
starting from the baseline observed in the data.

### 💰 Revenue Sensitivity

```{r financial-projections}
completed <- transactions %>% filter(status == "Completed")
months_observed <- as.numeric(difftime(max(completed$transaction_date),
                                       min(completed$transaction_date),
                                       units = "days")) / (365.25 / 12)
current_metrics <- list(
  monthly_revenue = sum(completed$total_amount) / months_observed,
  avg_clv = mean(clv_data$historical_clv, na.rm = TRUE),
  avg_order_value = mean(completed$total_amount)
)

# ASSUMPTIONS, not estimates: relative uplift in revenue per customer
assumed_uplift <- c(0, 0.02, 0.05, 0.10, 0.15)
scenario_table <- tibble(
  `Assumed uplift` = scales::percent(assumed_uplift),
  `Monthly revenue` = scales::dollar(current_metrics$monthly_revenue * (1 + assumed_uplift)),
  `Additional annual revenue` = scales::dollar(12 * current_metrics$monthly_revenue * assumed_uplift)
)

scenario_table %>%
  kbl(caption = sprintf("Revenue under assumed uplifts (baseline: %s per month over %.1f observed months)",
                        scales::dollar(current_metrics$monthly_revenue), months_observed)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

### 📈 Revenue Growth Scenarios

```{r revenue-projection}
# Scenario paths under ASSUMED monthly growth rates (0%, 1%, 2%)
months <- 1:24
base_revenue <- current_metrics$monthly_revenue
projection_df <- expand.grid(month = months, growth = c(0, 0.01, 0.02)) %>%
  mutate(revenue = base_revenue * (1 + growth)^month,
         scenario = factor(scales::percent(growth), levels = c("0%", "1%", "2%")))

ggplot(projection_df, aes(x = month, y = revenue, color = scenario)) +
  geom_line(linewidth = 1.2) +
  scale_color_manual(values = c("#95A5A6", "#3498DB", "#27AE60")) +
  scale_y_continuous(labels = dollar_format()) +
  labs(title = "24-Month Revenue Paths Under Assumed Growth Rates",
       subtitle = "Illustrative scenarios from the observed baseline; growth rates are assumptions",
       x = "Month", y = "Monthly Revenue", color = "Assumed monthly growth") +
  theme_professional() +
  theme(legend.position = "bottom")
```

---

# Implementation Framework {.tabset}

## Technology Stack

### 🛠️ Required Technology Infrastructure

```{r tech-stack}
# Create technology requirements table
tech_requirements <- tribble(
  ~Category, ~Technology, ~Purpose, ~Priority,
  "Analytics Platform", "R/RStudio Server", "Advanced analytics and modeling", "High",
  "Data Warehouse", "PostgreSQL/BigQuery", "Centralized data storage", "High",
  "BI Dashboard", "Shiny/Tableau", "Executive reporting and KPIs", "High",
  "Customer Data Platform", "Segment/mParticle", "Unified customer profiles", "Medium",
  "Email Marketing", "Mailchimp/Klaviyo", "Automated campaigns", "Medium",
  "A/B Testing", "Optimizely/VWO", "Campaign optimization", "Medium",
  "ML Operations", "MLflow/Kubeflow", "Model deployment and monitoring", "Low"
)

tech_requirements %>%
  kbl(caption = "Technology Infrastructure Requirements") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  column_spec(4, background = "#e8f4fd") %>%
  pack_rows("Core Analytics (Year 1)", 1, 3, background = "#d4edda") %>%
  pack_rows("Customer Experience (Year 2)", 4, 6, background = "#e8f4fd") %>%
  pack_rows("Advanced Operations (Year 3)", 7, 7, background = "#f8d7da")
```

## Performance Metrics

### 📊 Key Performance Indicators (KPIs)

```{r kpi-framework}
# Define comprehensive KPI framework
kpi_framework <- tribble(
  ~Category, ~KPI, ~Current_Value, ~Measurement_Frequency, ~Owner,
  "Revenue", "Monthly Revenue", scales::dollar(current_metrics$monthly_revenue), "Monthly", "CFO",
  "Revenue", "Average Order Value", scales::dollar(current_metrics$avg_order_value), "Weekly", "VP Sales",
  "Customer", "Average Historical CLV", scales::dollar(current_metrics$avg_clv), "Quarterly", "CMO",
  "Customer", "Share of High-Risk Customers",
    scales::percent(mean(churn_data$churn_risk_level == "High Risk"), accuracy = 0.1), "Quarterly", "Customer Success",
  "Customer", "Average Satisfaction (1-10)",
    sprintf("%.2f", mean(customers$satisfaction_score, na.rm = TRUE)), "Quarterly", "Customer Success",
  "Operational", "Order Completion Rate",
    scales::percent(mean(transactions$status == "Completed"), accuracy = 0.1), "Daily", "COO"
)

kpi_framework %>%
  kbl(caption = "KPI Framework: current values computed from the (simulated) data; targets are for stakeholders to set") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  pack_rows("Financial Metrics", 1, 2, background = "#fff3cd") %>%
  pack_rows("Customer Metrics", 3, 5, background = "#e8f4fd") %>%
  pack_rows("Operational Metrics", 6, 6, background = "#f8d7da")
```

### 🎯 Success Measurement Framework

```{r success-framework}
# Create success measurement timeline
success_timeline <- tribble(
  ~Timeframe, ~Milestone, ~Success_Criteria, ~Measurement_Method,
  "Month 1-3", "Foundation Setup", "Data pipeline operational, KPI dashboard live", "Technical validation, stakeholder sign-off",
  "Month 4-6", "Initial Campaigns", "Pre-registered effect on target KPI vs. hold-out", "Randomised A/B tests",
  "Month 7-9", "Scale & Optimize", "Effect replicated in a second cohort", "Cohort analysis, customer feedback",
  "Month 10-12", "Full Implementation", "Sustained effect at full rollout", "Year-over-year comparison, cost-benefit analysis",
  "Month 13+", "Continuous Improvement", "Sustained growth, new opportunity identification", "Ongoing optimization, strategic planning"
)

success_timeline %>%
  kbl(caption = "Success Measurement Timeline") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  column_spec(1, background = "#e8f4fd", bold = TRUE) %>%
  column_spec(3, background = "#d4edda")
```

---

# Conclusion

## Key Takeaways

```{r conclusion-metrics, include = FALSE}
seg_rev <- completed %>%
  left_join(customers %>% select(customer_id, customer_segment), by = "customer_id") %>%
  group_by(customer_segment) %>%
  summarise(revenue_per_customer = sum(total_amount) / n_distinct(customer_id), .groups = "drop")
rev_ratio <- with(seg_rev, max(revenue_per_customer) / min(revenue_per_customer))
top_seg <- seg_rev$customer_segment[which.max(seg_rev$revenue_per_customer)]
low_seg <- seg_rev$customer_segment[which.min(seg_rev$revenue_per_customer)]
q4_revenue_share <- sum(completed$total_amount[lubridate::quarter(completed$transaction_date) == 4]) /
  sum(completed$total_amount)
```

On the simulated data:

1. **Segments**: k-means on standardised behaviour and demographics found
   `r k` clusters (average silhouette width `r round(max(avg_silhouette), 3)`).
   A silhouette width below about 0.5 means the clusters are weakly
   separated and should not be over-interpreted.
2. **Revenue concentration**: `r top_seg` customers generate
   `r sprintf("%.1f", rev_ratio)` times as much revenue per customer as
   `r low_seg` customers.
3. **Churn risk**: `r scales::percent(mean(churn_data$churn_risk_level == "High Risk"), accuracy = 0.1)`
   of customers with purchases are classed as high risk by the rule-based
   score. The score is heuristic, so its thresholds should be validated
   against observed churn before being used.
4. **Seasonality**: Q4 accounts for `r scales::percent(q4_revenue_share, accuracy = 0.1)`
   of revenue, against 25% under a uniform spread.

Financial effects of the proposed initiatives are **not** estimated here.
They must be measured with controlled experiments before being projected.

## Next Steps

1. **Immediate Actions (Next 30 days)**:
   - Set up data infrastructure and KPI dashboards
   - Begin high-risk customer retention campaigns
   - Implement basic customer segmentation in marketing systems

2. **Short-term Goals (3-6 months)**:
   - Launch premium customer loyalty program
   - Deploy automated email marketing campaigns
   - Establish A/B testing framework

3. **Long-term Vision (6-12 months)**:
   - Advanced predictive modeling for personalization
   - Real-time customer experience optimization
   - Expansion to new customer segments and markets

*All data in this case study are simulated; the analysis demonstrates methods rather than findings about a real business.*

---

*This analysis was conducted using R and represents a comprehensive approach to business intelligence and customer analytics. All code, methodologies, and visualizations are available for review and can be adapted for different business contexts.*