---
title: "Advanced Predictive Modeling Framework"
subtitle: "End-to-End Machine Learning Pipeline for Business Applications"
author: "R Data Science Portfolio"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: flatly
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: true
    toc_depth: 4
    number_sections: true
    code_folding: show
    df_print: paged
    fig_width: 12
    fig_height: 8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  warning = FALSE, 
  message = FALSE,
  fig.align = "center",
  cache = FALSE,
  dpi = 300
)

# Fixed analysis date so that results do not depend on the render date
analysis_date <- as.Date("2024-12-31")

# caret's "xgbTree" method is incompatible with xgboost >= 3.0; fall back to
# caret's "gbm" gradient boosting in that case.
use_xgb <- utils::packageVersion("xgboost") < "3.0.0"

# Load required libraries
library(dplyr)
library(ggplot2)
library(caret)
library(randomForest)
library(xgboost)
library(gbm)
library(themis)  # needed by caret for sampling = "smote"
library(glmnet)
library(e1071)
library(ROCR)
library(pROC)
library(corrplot)
library(VIM)
library(plotly)
library(kableExtra)
library(DT)
library(scales)
library(viridis)
library(lime)
library(DALEX)
library(modelr)
library(broom)

# Source custom functions
if (file.exists("../utils/data-generators.R")) {
  source("../utils/data-generators.R")
}
if (file.exists("../05-machine-learning/supervised-learning.R")) {
  source("../05-machine-learning/supervised-learning.R")
}
if (file.exists("../05-machine-learning/feature-engineering.R")) {
  source("../05-machine-learning/feature-engineering.R")
}
if (file.exists("../utils/plotting-helpers.R")) {
  source("../utils/plotting-helpers.R")
}

# Set theme
theme_set(theme_minimal() + 
          theme(plot.title = element_text(size = 14, face = "bold"),
                plot.subtitle = element_text(size = 12, color = "gray40")))
```

# Executive Summary

This comprehensive predictive modeling analysis demonstrates advanced machine learning techniques using R. We develop end-to-end prediction models for customer behavior, implementing industry best practices for feature engineering, model selection, evaluation, and deployment.

## Project Objectives

- **Primary Goal**: Develop robust predictive models for customer lifetime value and churn prediction
- **Secondary Goals**: Demonstrate feature engineering, model comparison, and interpretability
- **Business Impact**: Enable data-driven customer retention and revenue optimization strategies

## Key Results

All data in this case study are **simulated** (see `modules/utils/data-generators.R`),
so the results illustrate methods rather than a real business. The model
comparison and every performance figure are computed below; see the
[technical appendix](#technical-appendix) for the summary table.

---

# Data Preparation and Feature Engineering

## Dataset Generation and Overview

```{r data-generation}
# Generate comprehensive business dataset for modeling
set.seed(42)

# Create realistic e-commerce dataset with target variables
modeling_data <- generate_comprehensive_business_data(
  n_customers = 3000,
  n_products = 250,
  n_transactions = 8000,
  date_range = c(as.Date("2021-01-01"), as.Date("2024-12-31"))
)

# Extract and combine datasets
customers <- modeling_data$customers
transactions <- modeling_data$transactions
customer_metrics <- modeling_data$customer_metrics

# Create comprehensive modeling dataset
analysis_data <- customers %>%
  left_join(customer_metrics, by = "customer_id") %>%
  # Customers without completed orders have spent nothing (not an unknown
  # amount); treat their registration as their most recent activity.
  mutate(
    total_orders = coalesce(total_orders, 0L),
    total_spent = coalesce(total_spent, 0),
    purchase_frequency = coalesce(purchase_frequency, 0),
    last_purchase = coalesce(last_purchase, registration_date)
  ) %>%
  mutate(
    # Target Variables for Different Problems
    
    # 1. Customer Lifetime Value (Regression)
    customer_ltv = total_spent * 1.5 + purchase_frequency * 200 + 
                   ifelse(customer_segment == "Premium", 500, 0) +
                   rnorm(n(), 0, 100),
    
    # 2. Churn Prediction (Binary Classification)
    days_since_last = as.numeric(analysis_date - last_purchase),
    churn_risk_score = pmax(0, pmin(1, 
      0.3 * (days_since_last / 365) + 
      0.2 * (1 - satisfaction_score / 10) + 
      0.1 * ifelse(purchase_frequency < 1, 1, 0) +
      rnorm(n(), 0, 0.1)
    )),
    will_churn = ifelse(churn_risk_score > 0.6, 1, 0),
    
    # 3. Segment Prediction (Multi-class Classification)
    segment_numeric = as.numeric(as.factor(customer_segment)),
    
    # Feature Engineering
    customer_age_days = as.numeric(analysis_date - registration_date),
    spending_per_day = total_spent / pmax(1, customer_age_days),
    satisfaction_category = cut(satisfaction_score, 
                               breaks = c(0, 5, 7, 10), 
                               labels = c("Low", "Medium", "High")),
    high_value_customer = ifelse(total_spent > quantile(total_spent, 0.8), 1, 0),
    
    # Interaction features
    age_income_interaction = age * annual_income / 1000000,
    frequency_satisfaction = purchase_frequency * satisfaction_score
  )

# Display dataset overview
cat("Modeling Dataset Overview:")
cat("\n- Total Customers:", nrow(analysis_data))
cat("\n- Total Features:", ncol(analysis_data))
cat("\n- Target Variables: CLV (continuous), Churn (binary), Segment (multi-class)")
```

## Advanced Feature Engineering

```{r feature-engineering}
# Apply comprehensive feature engineering
engineered_data <- analysis_data %>%
  # Create polynomial features
  mutate(
    age_squared = age^2,
    income_log = log(annual_income + 1),
    spending_sqrt = sqrt(total_spent + 1)
  ) %>%
  # Create binned features
  mutate(
    age_group = cut(age, breaks = c(0, 25, 35, 50, 65, 100), 
                   labels = c("18-25", "26-35", "36-50", "51-65", "65+")),
    income_quartile = ntile(annual_income, 4),
    spending_decile = ntile(total_spent, 10)
  ) %>%
  # Create time-based features
  mutate(
    registration_year = year(registration_date),
    registration_month = month(registration_date),
    tenure_years = as.numeric(analysis_date - registration_date) / 365.25,
    frequent_north_american = ifelse(purchase_frequency > median(purchase_frequency) & 
                                country %in% c("USA", "Canada"), 1, 0)
  )

# Feature importance preview
numeric_features <- engineered_data %>% 
  select_if(is.numeric) %>%
  select(-any_of(c("customer_id", "segment_numeric"))) %>%
  names()

cat("Engineered Features Summary:")
cat("\n- Original features:", ncol(analysis_data))
cat("\n- Engineered features:", ncol(engineered_data))  
cat("\n- Numeric features for modeling:", length(numeric_features))
```

### Feature Distribution Analysis

```{r feature-distribution, fig.height=10}
# Analyze key feature distributions
key_features <- c("customer_ltv", "age", "annual_income", "total_spent", 
                 "purchase_frequency", "satisfaction_score")

# Create distribution plots
distribution_plots <- list()

for (i in seq_along(key_features)) {
  feature <- key_features[i]
  
  if (feature %in% names(engineered_data)) {
    p <- ggplot(engineered_data, aes_string(x = feature)) +
      geom_histogram(bins = 30, fill = "#3498DB", alpha = 0.7, color = "white") +
      geom_density(aes(y = ..density.. * nrow(engineered_data) * 
                      (max(engineered_data[[feature]], na.rm = TRUE) - 
                       min(engineered_data[[feature]], na.rm = TRUE)) / 30),
                  color = "#E74C3C", size = 1) +
      labs(title = paste("Distribution of", tools::toTitleCase(gsub("_", " ", feature))),
           x = tools::toTitleCase(gsub("_", " ", feature)), y = "Count") +
      theme_professional()
    
    # Format currency/percentage labels
    if (grepl("ltv|income|spent", feature)) {
      p <- p + scale_x_continuous(labels = dollar_format())
    }
    
    distribution_plots[[i]] <- p
  }
}

# Arrange distribution plots
do.call(gridExtra::grid.arrange, c(distribution_plots, ncol = 3))
```

---

# Model Development Framework

## Problem 1: Customer Lifetime Value Prediction (Regression)

### Data Preparation for CLV Model

```{r clv-data-prep}
# Prepare CLV prediction dataset
clv_data <- engineered_data %>%
  select(customer_ltv, age, annual_income, total_spent, purchase_frequency, 
         satisfaction_score, tenure_years, age_squared, income_log, 
         spending_sqrt, age_income_interaction, frequency_satisfaction) %>%
  filter(complete.cases(.))

# Split data
set.seed(42)
train_indices <- createDataPartition(clv_data$customer_ltv, p = 0.7, list = FALSE)
clv_train <- clv_data[train_indices, ]
clv_test <- clv_data[-train_indices, ]

cat("CLV Dataset:")
cat("\n- Training samples:", nrow(clv_train))
cat("\n- Test samples:", nrow(clv_test))
cat("\n- Features:", ncol(clv_data) - 1)
```

### CLV Model Training and Comparison

```{r clv-models, results='hide'}
# Train multiple regression models
clv_models <- list()

# 1. Linear Regression
clv_models$linear <- train(
  customer_ltv ~ ., 
  data = clv_train,
  method = "lm",
  trControl = trainControl(method = "cv", number = 5)
)

# 2. Random Forest
clv_models$rf <- train(
  customer_ltv ~ ., 
  data = clv_train,
  method = "rf",
  trControl = trainControl(method = "cv", number = 5),
  tuneGrid = expand.grid(mtry = c(2, 4, 6)),
  ntree = 100
)

# 3. Gradient boosting (xgboost, or gbm with xgboost >= 3.0)
if (use_xgb) {
  clv_models$xgb <- train(
    customer_ltv ~ ., 
    data = clv_train,
    method = "xgbTree",
    trControl = trainControl(method = "cv", number = 5),
    tuneGrid = expand.grid(
      nrounds = c(50, 100),
      max_depth = c(3, 6),
      eta = c(0.1, 0.3),
      gamma = 0,
      colsample_bytree = 1,
      min_child_weight = 1,
      subsample = 1
    ),
    verbose = FALSE
  )
} else {
  clv_models$gbm <- train(
    customer_ltv ~ ., 
    data = clv_train,
    method = "gbm",
    trControl = trainControl(method = "cv", number = 5),
    tuneGrid = expand.grid(
      n.trees = c(50, 100),
      interaction.depth = c(3, 6),
      shrinkage = c(0.1, 0.3),
      n.minobsinnode = 10
    ),
    verbose = FALSE
  )
}

# 4. Ridge Regression
clv_models$ridge <- train(
  customer_ltv ~ ., 
  data = clv_train,
  method = "glmnet",
  trControl = trainControl(method = "cv", number = 5),
  tuneGrid = expand.grid(alpha = 0, lambda = seq(0.01, 1, by = 0.1))
)
```

### CLV Model Evaluation

```{r clv-evaluation}
# Generate predictions
clv_predictions <- map_dfr(names(clv_models), function(model_name) {
  model <- clv_models[[model_name]]
  pred <- predict(model, clv_test)
  
  data.frame(
    model = model_name,
    actual = clv_test$customer_ltv,
    predicted = pred
  )
})

# Calculate performance metrics
clv_performance <- clv_predictions %>%
  group_by(model) %>%
  summarise(
    RMSE = sqrt(mean((actual - predicted)^2)),
    MAE = mean(abs(actual - predicted)),
    # Out-of-sample R² (1 - SSE/SST); cor()^2 would ignore bias and scale errors
    R_squared = 1 - sum((actual - predicted)^2) / sum((actual - mean(actual))^2),
    .groups = "drop"
  ) %>%
  arrange(RMSE)

# Display performance table
clv_performance %>%
  mutate(
    RMSE = round(RMSE, 2),
    MAE = round(MAE, 2),
    R_squared = round(R_squared, 3)
  ) %>%
  kbl(caption = "Customer Lifetime Value Model Performance") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  row_spec(1, bold = TRUE, background = "#d4edda")
```

### CLV Model Visualization

```{r clv-visualization, fig.height=10}
# Model performance visualization
# Choose the model by its cross-validated RMSE on the training data. Picking
# the lowest *test* RMSE and then reporting it would be optimistically biased.
cv_rmse <- sapply(clv_models, function(m) min(m$results$RMSE))
best_model_name <- names(which.min(cv_rmse))
best_model_data <- clv_predictions %>% filter(model == best_model_name)

# 1. Prediction vs Actual
p1 <- ggplot(best_model_data, aes(x = actual, y = predicted)) +
  geom_point(alpha = 0.6, color = "#3498DB") +
  geom_abline(intercept = 0, slope = 1, color = "#E74C3C", linetype = "dashed") +
  geom_smooth(method = "lm", se = TRUE, color = "#27AE60") +
  scale_x_continuous(labels = dollar_format()) +
  scale_y_continuous(labels = dollar_format()) +
  labs(title = paste("CLV Predictions:", toupper(best_model_name)),
       subtitle = paste("R² =", round(clv_performance$R_squared[1], 3)),
       x = "Actual CLV", y = "Predicted CLV") +
  theme_professional()

# 2. Residuals plot
residuals_data <- best_model_data %>%
  mutate(residuals = actual - predicted)

p2 <- ggplot(residuals_data, aes(x = predicted, y = residuals)) +
  geom_point(alpha = 0.6, color = "#3498DB") +
  geom_hline(yintercept = 0, color = "#E74C3C", linetype = "dashed") +
  geom_smooth(method = "loess", se = TRUE, color = "#F39C12") +
  scale_x_continuous(labels = dollar_format()) +
  scale_y_continuous(labels = dollar_format()) +
  labs(title = "Residuals Analysis",
       x = "Predicted CLV", y = "Residuals") +
  theme_professional()

# 3. Model comparison
p3 <- ggplot(clv_performance, aes(x = reorder(model, R_squared), y = R_squared)) +
  geom_col(fill = "#9B59B6", alpha = 0.8) +
  geom_text(aes(label = round(R_squared, 3)), vjust = -0.5) +
  coord_flip() +
  labs(title = "Model Performance Comparison (R²)",
       x = "Model", y = "R-squared") +
  theme_professional()

# 4. Feature importance (for best model)
if (best_model_name %in% c("rf", "xgb", "gbm")) {
  importance_data <- varImp(clv_models[[best_model_name]])$importance
  importance_df <- data.frame(
    feature = rownames(importance_data),
    importance = importance_data[, 1]
  ) %>%
    arrange(desc(importance)) %>%
    head(10)
  
  p4 <- ggplot(importance_df, aes(x = reorder(feature, importance), y = importance)) +
    geom_col(fill = "#1ABC9C", alpha = 0.8) +
    coord_flip() +
    labs(title = "Top 10 Feature Importance",
         x = "Features", y = "Importance") +
    theme_professional()
} else {
  p4 <- ggplot() + 
    labs(title = "Feature importance not available for linear models") +
    theme_void()
}

# Arrange plots
gridExtra::grid.arrange(p1, p2, p3, p4, ncol = 2)
```

---

## Problem 2: Churn Prediction (Binary Classification)

### Churn Data Preparation

```{r churn-data-prep}
# Prepare churn prediction dataset
churn_data <- engineered_data %>%
  select(will_churn, age, annual_income, total_spent, purchase_frequency, 
         satisfaction_score, days_since_last, tenure_years, high_value_customer,
         age_squared, income_log, frequency_satisfaction) %>%
  filter(complete.cases(.)) %>%
  mutate(will_churn = factor(will_churn, levels = c(0, 1), labels = c("Retain", "Churn")))

# Check class balance
churn_balance <- table(churn_data$will_churn)
cat("Churn Dataset Class Distribution:")
print(churn_balance)
cat("Class proportions:", round(prop.table(churn_balance), 3))

# Split data
set.seed(42)
train_indices <- createDataPartition(churn_data$will_churn, p = 0.7, list = FALSE)
churn_train <- churn_data[train_indices, ]
churn_test <- churn_data[-train_indices, ]
```

### Churn Model Training

```{r churn-models, results='hide'}
# Train classification models
churn_models <- list()

# Control with class balancing
ctrl <- trainControl(
  method = "cv", 
  number = 5,
  classProbs = TRUE,
  summaryFunction = twoClassSummary,
  sampling = "smote"  # Handle class imbalance
)

# 1. Logistic Regression
churn_models$logistic <- train(
  will_churn ~ ., 
  data = churn_train,
  method = "glm",
  family = "binomial",
  trControl = ctrl,
  metric = "ROC"
)

# 2. Random Forest
churn_models$rf <- train(
  will_churn ~ ., 
  data = churn_train,
  method = "rf",
  trControl = ctrl,
  tuneGrid = expand.grid(mtry = c(2, 4, 6)),
  metric = "ROC",
  ntree = 100
)

# 3. SVM
churn_models$svm <- train(
  will_churn ~ ., 
  data = churn_train,
  method = "svmRadial",
  trControl = ctrl,
  tuneGrid = expand.grid(C = c(0.1, 1, 10), sigma = c(0.01, 0.1, 1)),
  metric = "ROC"
)

# 4. Gradient boosting (xgboost, or gbm with xgboost >= 3.0)
if (use_xgb) {
  churn_models$xgb <- train(
    will_churn ~ ., 
    data = churn_train,
    method = "xgbTree",
    trControl = ctrl,
    tuneGrid = expand.grid(
      nrounds = c(50, 100),
      max_depth = c(3, 6),
      eta = c(0.1, 0.3),
      gamma = 0,
      colsample_bytree = 1,
      min_child_weight = 1,
      subsample = 1
    ),
    metric = "ROC",
    verbose = FALSE
  )
} else {
  churn_models$gbm <- train(
    will_churn ~ ., 
    data = churn_train,
    method = "gbm",
    trControl = ctrl,
    tuneGrid = expand.grid(
      n.trees = c(50, 100),
      interaction.depth = c(3, 6),
      shrinkage = c(0.1, 0.3),
      n.minobsinnode = 10
    ),
    metric = "ROC",
    verbose = FALSE
  )
}
```

### Churn Model Evaluation

```{r churn-evaluation}
# Generate predictions and probabilities
churn_predictions <- map_dfr(names(churn_models), function(model_name) {
  model <- churn_models[[model_name]]
  pred_class <- predict(model, churn_test)
  pred_prob <- predict(model, churn_test, type = "prob")
  
  data.frame(
    model = model_name,
    actual = churn_test$will_churn,
    predicted = pred_class,
    prob_churn = pred_prob$Churn
  )
})

# Calculate performance metrics
churn_performance <- churn_predictions %>%
  group_by(model) %>%
  summarise(
    Accuracy = mean(actual == predicted),
    Precision = sum(actual == "Churn" & predicted == "Churn") / sum(predicted == "Churn"),
    Recall = sum(actual == "Churn" & predicted == "Churn") / sum(actual == "Churn"),
    F1_Score = 2 * (Precision * Recall) / (Precision + Recall),
    AUC = as.numeric(pROC::auc(pROC::roc(as.numeric(actual == "Churn"), prob_churn, quiet = TRUE))),
    .groups = "drop"
  ) %>%
  arrange(desc(AUC))

# Display performance table
churn_performance %>%
  mutate_if(is.numeric, ~round(.x, 3)) %>%
  kbl(caption = "Churn Prediction Model Performance") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  row_spec(1, bold = TRUE, background = "#d4edda")
```

### Churn Model Visualization

```{r churn-visualization, fig.height=12}
# ROC Curves for all models
roc_data <- map_dfr(names(churn_models), function(model_name) {
  model_preds <- churn_predictions %>% filter(model == model_name)
  roc_obj <- pROC::roc(as.numeric(model_preds$actual == "Churn"), model_preds$prob_churn, quiet = TRUE)
  
  data.frame(
    model = model_name,
    specificity = roc_obj$specificities,
    sensitivity = roc_obj$sensitivities,
    auc = as.numeric(pROC::auc(roc_obj))
  )
})

# 1. ROC Curves
p1 <- ggplot(roc_data, aes(x = 1 - specificity, y = sensitivity, color = model)) +
  geom_line(size = 1.2) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray") +
  scale_color_manual(values = get_custom_palette("professional", 4)) +
  labs(title = "ROC Curves for Churn Prediction Models",
       x = "1 - Specificity (False Positive Rate)",
       y = "Sensitivity (True Positive Rate)",
       color = "Model") +
  theme_professional()

# 2. Precision-Recall curves
pr_data <- map_dfr(names(churn_models), function(model_name) {
  model_preds <- churn_predictions %>% filter(model == model_name)
  
  # Calculate precision-recall points
  thresholds <- seq(0, 1, by = 0.01)
  pr_points <- map_dfr(thresholds, function(threshold) {
    pred_binary <- ifelse(model_preds$prob_churn > threshold, "Churn", "Retain")
    
    if (sum(pred_binary == "Churn") > 0) {
      precision <- sum(model_preds$actual == "Churn" & pred_binary == "Churn") / 
                   sum(pred_binary == "Churn")
      recall <- sum(model_preds$actual == "Churn" & pred_binary == "Churn") / 
                sum(model_preds$actual == "Churn")
    } else {
      precision <- 1
      recall <- 0
    }
    
    data.frame(threshold = threshold, precision = precision, recall = recall)
  })
  
  pr_points$model <- model_name
  return(pr_points)
})

p2 <- ggplot(pr_data, aes(x = recall, y = precision, color = model)) +
  geom_line(size = 1.2) +
  scale_color_manual(values = get_custom_palette("professional", 4)) +
  labs(title = "Precision-Recall Curves",
       x = "Recall", y = "Precision", color = "Model") +
  theme_professional()

# 3. Confusion Matrix for best model
# Choose by cross-validated ROC AUC; the test set is only used to report it.
cv_auc <- sapply(churn_models, function(m) max(m$results$ROC))
best_churn_model <- names(which.max(cv_auc))
best_churn_preds <- churn_predictions %>% filter(model == best_churn_model)

conf_matrix <- table(Predicted = best_churn_preds$predicted, 
                    Actual = best_churn_preds$actual)

conf_df <- as.data.frame(conf_matrix) %>%
  mutate(
    Percentage = round(Freq / sum(Freq) * 100, 1),
    Label = paste0(Freq, "\n(", Percentage, "%)")
  )

p3 <- ggplot(conf_df, aes(x = Actual, y = Predicted, fill = Freq)) +
  geom_tile(alpha = 0.8) +
  geom_text(aes(label = Label), size = 4, color = "black") +
  scale_fill_gradient(low = "white", high = "#3498DB") +
  labs(title = paste("Confusion Matrix:", toupper(best_churn_model)),
       subtitle = paste("Accuracy:", round(churn_performance$Accuracy[churn_performance$model == best_churn_model], 3))) +
  theme_professional() +
  theme(legend.position = "none")

# 4. Feature importance for best model
if (best_churn_model %in% c("rf", "xgb", "gbm")) {
  importance_data <- varImp(churn_models[[best_churn_model]])$importance
  importance_df <- data.frame(
    feature = rownames(importance_data),
    importance = importance_data[, 1]
  ) %>%
    arrange(desc(importance)) %>%
    head(10)
  
  p4 <- ggplot(importance_df, aes(x = reorder(feature, importance), y = importance)) +
    geom_col(fill = "#E74C3C", alpha = 0.8) +
    coord_flip() +
    labs(title = "Top 10 Features for Churn Prediction",
         x = "Features", y = "Importance") +
    theme_professional()
} else {
  p4 <- ggplot() + 
    labs(title = "Feature importance not available for this model") +
    theme_void()
}

# Arrange plots
gridExtra::grid.arrange(p1, p2, p3, p4, ncol = 2)
```

---

## Problem 3: Customer Segment Classification (Multi-class)

### Segment Classification Setup

```{r segment-classification}
# Prepare segment classification dataset
segment_data <- engineered_data %>%
  select(customer_segment, age, annual_income, total_spent, purchase_frequency, 
         satisfaction_score, tenure_years, high_value_customer,
         age_income_interaction, spending_per_day) %>%
  filter(complete.cases(.)) %>%
  mutate(customer_segment = factor(customer_segment))

# Check class distribution
segment_balance <- table(segment_data$customer_segment)
cat("Segment Classification Class Distribution:")
print(segment_balance)

# Split data
set.seed(42)
train_indices <- createDataPartition(segment_data$customer_segment, p = 0.7, list = FALSE)
segment_train <- segment_data[train_indices, ]
segment_test <- segment_data[-train_indices, ]

# Train best performing model (based on churn results)
segment_model <- train(
  customer_segment ~ ., 
  data = segment_train,
  method = "rf",
  trControl = trainControl(method = "cv", number = 5),
  tuneGrid = expand.grid(mtry = c(2, 4, 6)),
  ntree = 100
)

# Predictions
segment_pred <- predict(segment_model, segment_test)
segment_accuracy <- mean(segment_pred == segment_test$customer_segment)

# Compare with the no-information rate (always predicting the largest class)
segment_nir <- max(prop.table(table(segment_test$customer_segment)))

cat("\nSegment Classification Results:")
cat("\n- Model: Random Forest")
cat("\n- Accuracy:", round(segment_accuracy, 3))
cat("\n- No-information rate:", round(segment_nir, 3))
```

---

# Model Interpretability and Explainability

## LIME Explanations

```{r model-interpretability, fig.height=8}
# LIME explanations for the best churn model
if (requireNamespace("lime", quietly = TRUE)) {
  
  # Prepare data for LIME
  churn_features <- churn_test %>% select(-will_churn)
  
  # Create explainer
  # DALEX also exports explain(), so call lime's functions explicitly
  explainer <- lime::lime(churn_train %>% select(-will_churn), 
                   churn_models[[best_churn_model]])
  
  # Explain a few predictions
  explanations <- lime::explain(churn_features[1:4, ], explainer, n_labels = 1,
                                n_features = 5)
  
  # Plot explanations
  lime::plot_features(explanations) +
    labs(title = "LIME Explanations for Churn Predictions") +
    theme_professional()
}
```

## Global Feature Importance

```{r global-importance, fig.height=6}
# Scaled (0-100) variable importance of the selected CLV and churn models
importance_of <- function(model, label) {
  imp <- varImp(model)$importance
  data.frame(Feature = rownames(imp), Importance = imp[, 1], Problem = label)
}
importance_comparison <- rbind(
  importance_of(clv_models[[best_model_name]], "CLV"),
  importance_of(churn_models[[best_churn_model]], "Churn"),
  importance_of(segment_model, "Segment")
)
top_features <- importance_comparison %>%
  group_by(Feature) %>%
  summarise(max_imp = max(Importance), .groups = "drop") %>%
  slice_max(max_imp, n = 8) %>%
  pull(Feature)

ggplot(filter(importance_comparison, Feature %in% top_features),
       aes(x = reorder(Feature, Importance), y = Importance, fill = Problem)) +
  geom_col(position = "dodge", alpha = 0.8) +
  coord_flip() +
  scale_fill_manual(values = get_custom_palette("professional", 3)) +
  labs(title = "Feature Importance of the Selected Models",
       subtitle = "caret::varImp(), scaled to 0-100 within each model",
       x = "Features", y = "Scaled importance", fill = "Problem") +
  theme_professional()
```

---

# Business Impact Analysis

## Model Performance on the Held-out Test Sets

The table reports test-set performance of the model selected for each problem
by cross-validation. Translating these into monetary value would require real
costs and intervention effects, which simulated data cannot supply, so no
financial projections are made here.

```{r business-impact}
clv_best <- clv_performance[clv_performance$model == best_model_name, ]
churn_best <- churn_performance[churn_performance$model == best_churn_model, ]
business_impact <- data.frame(
  Problem = c("CLV prediction", "Churn prediction", "Segment classification"),
  Selected_Model = c(best_model_name, best_churn_model, "rf"),
  Test_Performance = c(
    sprintf("R² = %.3f, RMSE = %.1f", clv_best$R_squared, clv_best$RMSE),
    sprintf("AUC = %.3f, recall = %.3f", churn_best$AUC, churn_best$Recall),
    sprintf("Accuracy = %.3f (no-information rate %.3f)", segment_accuracy, segment_nir)
  )
)

business_impact %>%
  kbl(caption = "Test-set performance of the selected models (simulated data)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

## Implementation Roadmap

```{r implementation-roadmap}
implementation_plan <- data.frame(
  Phase = c("Phase 1", "Phase 1", "Phase 2", "Phase 2", "Phase 3", "Phase 3"),
  Timeline = c("Month 1-2", "Month 1-2", "Month 3-4", "Month 3-4", "Month 5-6", "Month 5-6"),
  Activity = c("CLV Model Deployment", "Data Pipeline Setup", 
              "Churn Model Integration", "Real-time Scoring", 
              "Segment Model Launch", "Performance Monitoring"),
  Resources = c("2 Data Scientists", "1 ML Engineer", 
               "1 Data Scientist", "1 ML Engineer", 
               "1 Data Scientist", "1 Analytics Engineer"),
  Success_Metric = c("Model R² > 0.80", "Data latency < 1 hour",
                    "Model AUC > 0.85", "Prediction latency < 100ms",
                    "Model accuracy > 85%", "Model drift detection")
)

implementation_plan %>%
  kbl(caption = "Model Implementation Roadmap") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  pack_rows("Foundation Phase", 1, 2, background = "#fff3cd") %>%
  pack_rows("Core Implementation", 3, 4, background = "#e8f4fd") %>%
  pack_rows("Advanced Features", 5, 6, background = "#f8d7da")
```

---

# Model Monitoring and Maintenance

## Model Performance Tracking

```{r model-monitoring}
# Monitoring illustration built from the churn model's real test-set
# predictions: 24 weekly batches of 200 customers are resampled from the test
# set. With no drift at all, AUC still varies from week to week, and alert
# thresholds must allow for that sampling noise.
set.seed(42)
batch_auc <- function(idx) {
  as.numeric(pROC::auc(pROC::roc(best_churn_preds$actual[idx] == "Churn",
                                 best_churn_preds$prob_churn[idx], quiet = TRUE)))
}
monitoring_data <- data.frame(
  Week = 1:24,
  AUC = replicate(24, batch_auc(sample(nrow(best_churn_preds), 200, replace = TRUE)))
)
noise_band <- quantile(replicate(2000, batch_auc(sample(nrow(best_churn_preds), 200,
                                                         replace = TRUE))),
                       c(0.025, 0.975))

ggplot(monitoring_data, aes(x = Week, y = AUC)) +
  ggplot2::annotate("rect", xmin = -Inf, xmax = Inf, ymin = noise_band[1], ymax = noise_band[2],
           alpha = 0.15, fill = "#3498DB") +
  geom_hline(yintercept = churn_best$AUC, linetype = "dashed") +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  labs(title = "Churn Model Monitoring: Expected Week-to-Week Variation",
       subtitle = sprintf("Batches of 200 resampled from the test set; band = 95%% range without drift (%.3f to %.3f)",
                          noise_band[1], noise_band[2]),
       x = "Week", y = "Weekly AUC") +
  theme_professional()
```

## Deployment Considerations

### Model Deployment Checklist

```{r deployment-checklist}
deployment_checklist <- data.frame(
  Category = c("Data Quality", "Data Quality", "Model Performance", "Model Performance", 
              "Infrastructure", "Infrastructure", "Monitoring", "Monitoring"),
  Item = c("Input validation", "Feature drift detection", "Performance thresholds", 
          "A/B testing framework", "API endpoints", "Scalability testing",
          "Performance dashboards", "Alert systems"),
  Status = c("✅ Complete", "✅ Complete", "✅ Complete", "🔄 In Progress",
            "✅ Complete", "⚠️ Planned", "✅ Complete", "🔄 In Progress"),
  Priority = c("High", "High", "High", "Medium", "High", "Medium", "High", "High")
)

deployment_checklist %>%
  kbl(caption = "Model Deployment Readiness Checklist (template)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  column_spec(3, background = ifelse(grepl("Complete", deployment_checklist$Status), "#d4edda",
                             ifelse(grepl("Progress", deployment_checklist$Status), "#fff3cd", "#f8d7da")))
```

---

# Conclusions and Recommendations

## Key Achievements

### 🎯 **Model Performance** (test sets, simulated data)
1. **CLV Prediction**: the selected model (`r best_model_name`) reached an
   out-of-sample R² of `r round(clv_best$R_squared, 3)`. The simulated CLV is
   close to a deterministic function of spend and purchase frequency, which
   are themselves predictors, so a high R² is expected by construction.
2. **Churn Prediction**: `r best_churn_model` reached a test AUC of
   `r round(churn_best$AUC, 3)` and recall of `r round(churn_best$Recall, 3)`.
3. **Segment Classification**: accuracy was `r round(segment_accuracy, 3)`,
   compared with a no-information rate of `r round(segment_nir, 3)`.

### 💡 **Technical Insights**
1. **Model selection** used cross-validated performance on the training data
   only; the test set was used once, to report the selected model.
2. **Class imbalance** in the churn problem was handled with SMOTE inside the
   cross-validation folds (caret's `sampling = "smote"`), so synthetic cases
   never leak into the assessment folds.
3. **Interpretability**: LIME explains individual churn predictions, and
   `varImp()` summarises each selected model globally.

## Strategic Recommendations

### 🚀 **Immediate Actions (Next 30 days)**
1. **Deploy CLV Model**: Prioritize highest-impact model for revenue optimization
2. **Churn Prevention**: Implement real-time scoring for at-risk customer identification
3. **Data Pipeline**: Establish automated feature engineering and model scoring

### 🔧 **Medium-term Goals (3-6 months)**
1. **Model Enhancement**: Incorporate additional data sources (web behavior, support interactions)
2. **Real-time Inference**: Deploy low-latency prediction APIs for operational systems
3. **Advanced Segmentation**: Develop dynamic segmentation based on behavior patterns

### 📈 **Long-term Vision (6-12 months)**
1. **Personalization Engine**: Combine all models for individual customer optimization
2. **Automated Decision Making**: Implement ML-driven marketing and retention automation
3. **Predictive Analytics Platform**: Scale methodology across all business functions

---

# Technical Appendix

## Model Comparison Summary

```{r final-model-summary}
final_summary <- data.frame(
  Problem_Type = c("Regression", "Binary Classification", "Multi-class Classification"),
  Target_Variable = c("Customer Lifetime Value", "Churn", "Customer Segment"),
  Best_Algorithm = c(best_model_name, best_churn_model, "rf"),
  Selected_By = c("5-fold CV RMSE", "5-fold CV ROC AUC", "fixed (random forest)"),
  Test_Metric = business_impact$Test_Performance
)

final_summary %>%
  kbl(caption = "Final Model Selection and Performance Summary") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

## Session Information

```{r session-info}
sessionInfo()
```

---

*All data are simulated; the results illustrate the modelling workflow, not the performance achievable on real customer data.*